Extracting sunbeams from cucumbers: Reproducibility as a necessity for the future of archaeological science

### 2025-01-08 paper.qmd
intro

What has been missing from these debates is a macroscopic observation of what the majority of archaeologists are actually doing, and an objective comparison to a broad spectrum of relatively harder and softer disciplines.

>>>>

What has been missing from these debates is a macroscopic observation of what the majority of archaeologists are actually doing, and an empirical comparison to a broad spectrum of relatively harder and softer disciplines. At the 'hard' end of the spectrum, scholars more typically share an established set of theories, facts, and methods, facilitating fairly rapid agreement on the validity and significance of new results. At the 'soft' end of the spectrum, the theories, facts, and methods on which there is widespread consensus is smaller, and agreement is slower and less frequently reached about the significance of new finding and the continuing relevance of previous work. Hardness and softness is a controversial distinction, in part because it is often used to imply a rank order of disciplines that encodes legitimacy, productivity, perceived value to society, and worthiness of funding [@DifferentAgenda2012; @coleHierarchySciences1983]. Independent of these value judgments, empirical analysis of scholarly articles does indicate a spectrum of variation in practice linked to differing degrees of consensus in a discipline [@clevelandGraphsScientificPublications1984; @smithScientificGraphsHierarchy2000]. Furthermore, the hard-soft spectrum is evident in surveys of how researchers view their own work relative to others [@biglanCharacteristicsSubjectMatter1973].

n an ideal science, scholars share a common background of established theories, facts and methods. This allows them to agree (usually after debate and further evidence) on the validity and significance of a new research finding, making it the basis for further theorizing and research. Harder sciences are hypothesised to come closer to this ideal. Moving towards “softer” fields, this consensus becomes less likely to be reached, the common background shrinks and fractures, and so data become less able to “speak for themselves” [\[](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0066938#pone.0066938-Fanelli1)
